---
title: "BGC types in the A. nidulans assembly"
author: "Hosea"
date: "2026-10-03"
output: github_document
---

Cluster types from fungiSMASH (antiSMASH 8.0.4), split by whether each region
sits on a scaffold edge.

## Packages

```{r packages, message = FALSE}
library(ggplot2)
```

## Load the data

```{r load-data}
data_file <- "../08_antismash/antismash_regions.tsv"
file.exists(data_file)

regions <- read.delim(data_file, colClasses = c(label = "character"))
```

## Inspecting the data

```{r inspect}
str(regions)
```

## Tidy the data

```{r tidy}
regions$contig_edge <- as.logical(regions$contig_edge)

regions$type <- ifelse(grepl("+", regions$products, fixed = TRUE),
                       "Hybrid", regions$products)

regions$edge <- ifelse(regions$contig_edge, "On scaffold edge", "Not on edge")

type_order <- names(sort(table(regions$type)))
regions$type <- factor(regions$type, levels = type_order)

table(regions$type, regions$edge)
```

## Plot

```{r plot, fig.width = 7, fig.height = 4.5}
ggplot(regions, aes(x = type, fill = edge)) +
  geom_bar() +
  coord_flip() +
  scale_fill_manual(values = c("Not on edge" = "grey65",
                               "On scaffold edge" = "#D95F02")) +
  labs(x = NULL, y = "Number of regions", fill = NULL,
       title = "BGC regions by type",
       subtitle = "fungiSMASH, A. nidulans short-read assembly (66 regions)") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")
```

## Interpretation

PKS and NRPS regions were the most affected by assembly fragmentation:
9 of 17 T1PKS and 4 of 7 NRPS regions sit on a scaffold edge. Their core
genes are long, so these clusters are more likely to cross a break in a
short-read assembly.






